---
title: "Projections Blog"
author: "David Rios"
format:
  html:
    embed-resources: true
    
---

> “Every flat map misrepresents the surface of the Earth in some way.” - U.S. Department of the Interior


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

suppressPackageStartupMessages({
  library(tidyverse)
  library(sf)
  library(elevatr)
  library(terra)
  library(stars)
  library(tidycensus)
  library(USAboundaries)
})
```

## Introduction 

Hello, welcome to another blog. Today's topic is Projection.

At the heart of every spatial visualization is a set of locations. One way to describe a location is in terms of coordinates and a coordinate reference system (CRS). There are three main components to a CRS: ellipsoid, datum, and a projection. 
I’ll be focusing here on the last one, projection, but before we start, here’s a reminder of what those terms mean: 
    1. The ellipsoid is part of a CRS is a mathematical model giving a smooth approximation to Earth’s shape. 
    2. The center of the ellipsoid is called a datum. 
    3. projection is how we transform the Earth 3D dimension into a 2D graph. 
    
## How to choose map projections?

There’s a bunch of projections! For the more options refer to this cool [pdf](https://pubs.usgs.gov/gip/70047422/report.pdf) about map projections. Here are interesting ones: 

###  Mercator Projection
The one on top is called the Mercator projection from the 1500’s!  As you might be guessing, this is good for navigation as we have a standard north and south. The problem? It inflates some regions/areas, like the US.

```{r echo = FALSE}
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
# Basic Map w/ labels
ggplot(data = world) + 
    geom_sf(color = "black", fill = "#bada55") +
    labs(x = "Longitude", y = "Latitude", title = "World Map - Mercator Projection", subtitle = paste0("(", length(unique(world$name)), " countries)")) +
    theme_bw() 
```

### Lambert Azimuthal Equal-Area Projection

The second one was presented by Lambert in 1772 known as Lambert Azimuthal Equal-Area Projection. It is most Suited for regions extending equally in all directions from center points, such as Asia and Pacific Ocean. The problem? While it correctly represents the territory of each country, it does a bad job with the angles.

```{r echo = FALSE}
ggplot(data = world) +
    geom_sf(color = "black", fill = "#bada55") +
    coord_sf(crs = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs") + 
    labs(title = "Lambert Azimuthal Equal-Area Projection", subtitle = "Correctly represents area but not angles") + 
    theme_bw()
```


### Foucaut Projection

Lastly, I pick a funny one called Foucaut Projection or Sinusoidal Equal Area. In this projection, distances are correct along all parallels and the central meridian(s). The problem? Shapes are increasingly distorted away from the central meridian(s) and near the poles.

```{r echo = FALSE}
ggplot(data = world) +
    geom_sf(color = "black", fill = "#bada55") +
    coord_sf(crs = "+proj=fouc") + 
    labs(title = "Foucaut Projection", subtitle = "Correctly represents area, lots of shape distortion in high latitudes") + 
    theme_bw() 
```



> Question: I feel overwhelm, which one do I pick?

The answer will depend on your project, lol. But the main takeaway here is that you get to choose the projection you want to work with. The most used one, as stated above is the Mercator projection, but that does not mean you have to choose that one, right? 

I hope you have fun reading about projection. Thank you for reading this blog post. See you in another post.




