---
title: "W:10 How likely is 'likely'?"
format: html
---

## Introduction

People often use words like "likely," "probable," or "almost certain" to describe uncertain events, but how these phrases are interpreted can vary widely across individuals and cultures. In the *How Likely Is Likely?* dataset, over 5,000 participants assigned numerical probabilities (0–100%) to common probability phrases. This project explores how interpretation of these phrases differs across countries and demographic groups, highlighting global patterns of optimism and variation in probabilistic thinking.


## Research Question

> Which countries interpret probability phrases most optimistically?



## Data Import
```{r}
library(tidyverse)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)

tuesdata <- tidytuesdayR::tt_load(2026, week = 10)

absolute_judgements <- tuesdata$absolute_judgements
pairwise_comparisons <- tuesdata$pairwise_comparisons
respondent_metadata <- tuesdata$respondent_metadata
```

```{r final_plot}
# Merge probability estimates with metadata
data_full <- absolute_judgements |> 
  left_join(respondent_metadata, by = "response_id")

# Compute mean probability by country
country_summary <- data_full |>
  group_by(country_of_residence) |>
  summarize(
    mean_prob = mean(probability, na.rm = TRUE),
    n = n(),
    .groups = "drop")

# Fix country names to match rnaturalearth
country_summary <- country_summary |>
  mutate(country_of_residence = case_when(
    country_of_residence == "Federated States of Micronesia" ~ "Micronesia",
    country_of_residence == "Myanmar (Burma)" ~ "Myanmar",
    country_of_residence == "The Gambia" ~ "Gambia",
    country_of_residence == "United States" ~ "United States of America",
    TRUE ~ country_of_residence
  ))

# Load world map
world <- ne_countries(scale = "medium", returnclass = "sf")

# Merge world map with country_summary
world_data <- world |>
  left_join(country_summary, by = c("name" = "country_of_residence"))

# Plot world map
ggplot(world_data) +
  geom_sf(aes(fill = mean_prob)) +
  scale_fill_viridis_c(option = "C", na.value = "white", name = "Mean Probability") +
  labs(
    title = "Global Optimism in Interpreting Probability Phrases",
    subtitle = "Average probability assigned per country* (higher = more optimistic)",
    caption = "Data: Kucharski's 'How Likely Is Likely?' TidyTuesday 2026 W10. Author: David Rios\nWe have limited information on the countries responses; thus, multiple have no information.") +
  theme_classic() +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    plot.caption =element_text(hjust = 0))
```



