---
title: "W12: 10000 Pi Digits!"
format: html
format: 
  html: 
    embed-resources: true
---


## PI-DAY

In honour of pi day, this week TidyTuesday (TT) challenge is about the first 10000 digits of pi! 

In this TT challenge, I will be playing around with the pi digits, and plot them based on the golden ratio, a concept learned in my math classes. Then, I will paired each number with its corresponding color to see if any pattern emerge.

## Research Question

> Are the digits of pi (0–9) uniformly distributed across the first 10,000 decimal places?

## Data Preparation

```{r message=FALSE}
library(tidyverse)

tuesdata <- tidytuesdayR::tt_load(2026, week = 12)

pi_digits <- tuesdata$pi_digits
```

```{r pi_colors}

## some data preparation

golden_angle <- pi * (3 - sqrt(5))

df <- pi_digits %>%
  filter(digit_position <= 1000) %>%
  mutate(
    angle    = (digit_position - 1) * golden_angle,
    radius   = sqrt(digit_position - 1) * 1.9,
    x        = radius * cos(angle),
    y        = radius * sin(angle),
    dot_size = 0.3 + (radius / max(radius)) * 2.2,
    digit_f  = factor(digit, levels = 0:9))

# Color palette
pi_colors <- c(
  "0" = "#A0A0A0",   # silver
  "1" = "#FF4081",   # hot pink
  "2" = "#448AFF",   # blue
  "3" = "#FF6D00",   # orange
  "4" = "#00BCD4",   # cyan
  "5" = "#AA00FF",   # purple
  "6" = "#76FF03",   # lime green
  "7" = "#FFD740",   # amber
  "8" = "#1DE9B6",   # teal
  "9" = "#FF6E40"    # coral
)
```


## Final Plot, Enjoy! 

```{r pi_final_plot}

df |> 
ggplot(aes(x = x, y = y, colour = digit_f, size = dot_size)) +
  geom_point(shape = 16) +
  annotate("text", x = 0, y = 0, label = "\u03c0",
           colour = "white", size = 7, fontface = "bold") +
  scale_colour_manual(values = pi_colors) +
  scale_size_identity() +
  coord_fixed() +
  labs(
  title    = "10,000 Digits of Pi",
  subtitle = "Each dot represents a digit (0–9), coloured by its value",
  caption  = "Data: TidyTuesday Week 12 · Author: David Rios") +
theme_void() +
theme(
  plot.background  = element_rect(fill = "white", colour = NA),
  panel.background = element_rect(fill = "black", colour = NA),
  plot.title.position = "plot",
    legend.position = "none",   
  plot.title = element_text(
    family   = "Times New Roman",
    colour   = "grey10",
    size     = 22,
    face     = "bold",
    hjust    = 0.5,
    margin   = margin(t = 20, b = 6)
  ),
  plot.subtitle = element_text(
    family   = "Times New Roman",
    colour   = "grey30",
    size     = 13,
    hjust    = 0.5,
    margin   = margin(b = 20) ),
  plot.caption = element_text(
    family   = "Times New Roman",
    colour   = "grey50",
    hjust    = 0.5,
    margin   = margin(t = 16, b = 16)),
  plot.caption.position = "plot",
  plot.margin = margin(t = 10, r = 60, b = 10, l = 60))
```

